### 2011-04-05 docs/enthalpy/enthalpy.tex
So close ... don't want to lose this.

\section{Derivation of Constraint Using Enthalpy Equation}

The velocity constraint equation was constructed by differentiating 
the pressure along particle paths, using the equation of state.  
Specifically, we wrote $p_{EOS} = p(\rho,T,X_k) = p_0$
and expressed $DT/Dt$ in terms of $Dh/Dt:$
\begin{equation}
\frac{DT}{Dt} = \frac{1}{\rho c_p} \left( (1 - \rho h_p) \frac{D p}{D t}
- \sum_k \rho \xi_k \omegadot_k + \rho H \right) \enskip , \label{eq:dTdt}
\end{equation}
where $c_p = \left.\partial h/\partial T\right|_{p,X_k}$ is the
specific heat at constant pressure,
$\xi_k = \left.\partial h/\partial X_k \right|_{p,T}$,
and $h_p = \left.\partial h/\partial p\right|_{T,X_k}$
for convenience.

Then we derived
\begin{eqnarray*}
\nabla \cdot \ubold &=& \frac{1}{\rho p_\rho} \left(
- \frac{D p}{D t} + \frac{p_T}{\rho c_p}
  \left( (1 - \rho h_p) \frac{D p}{D t} - \rho \sum_k \xi_k \omegadot_k + \rho H \right)
+ \sum_k p_{X_k} \omegadot_k \right)  \\
                 &=& \frac{1}{\rho p_\rho}
  \left( \frac{p_T}{\rho c_p}(1  - \rho h_p) - 1 \right) \frac{D p}{D t}
 + \frac{1}{\rho p_\rho} \left(
  \frac{p_T}{\rho c_p} (\rho H - \rho \sum_k \xi_k   \omegadot_k)
                               + \sum_k p_{X_k} \omegadot_k \right) \enskip .
\end{eqnarray*}
When we derived this expression we explicitly retained the dependence of $h$ on $p$,
as shown by the presence of the $h_p$ term.

Then, replacing $p$ by $p_0(r)$, $Dp/Dt$ becomes $\ubold \cdot
\nabla p_0$, and, recalling the definition of $H$, the divergence constraint can
be written
\begin{equation}
\nabla \cdot \ubold + \alpha \ubold \cdot \nabla p_0 =
\frac{1}{\rho p_\rho} \left(
   \frac{p_T}{\rho c_p} \left(\nabla \cdot (\kappa \nabla T)
  - \sum_k\rho (q_k + \xi_k) \omegadot_k \right)
 + \sum_k p_{X_k} \omegadot_k \right)  \equiv \tilde{S} \enskip \label{eq:full_divu_constraint} ,
\end{equation}
where we define
\begin{equation}
\alpha(\rho,T) \equiv - \left( \frac{(1 - \rho h_p )p_T - \rho c_p}{\rho^2
  c_p p_\rho} \right) \enskip . \label{eq:alphadef}
\end{equation}

\section{Derivation of Constraint Using Energy Equation}

We could also derive the constraint by again differentiating $p_{EOS} = p(\rho,T,X_k) = p_0$
but this time writing $DT/Dt$ in terms of $De/Dt:$

\begin{equation}
\frac{DT}{Dt} = \frac{1}{\rho c_v} \left( -e_p \frac{D p}{D t}
- \sum_k e_k \frac{D X_k}{D t} - p \nabla \cdot U 
+ \rho H \right) \enskip , \label{eq:dTdt}
\end{equation}
where $c_v = \left.\partial e/\partial T\right|_{p,X_k}$ is the
specific heat at constant volume,
and $e_p = \left.\partial e/\partial p\right|_{T,X_k}$
for convenience.

Then...

>>>>

\section{Derivation of Velocity Constraint}

Differentiating the equation of state, written in the form, $p =
p(\rho,T,X_k),$ along particle paths, we can write
\begin{equation}
\frac{D p}{Dt}  = p_\rho \frac{D \rho}{Dt} + p_T \frac{D T}{Dt} + \sum_k p_{X_k} \frac{D X_k}{Dt}
\end{equation}
Then, by rearranging the terms, we get
\begin{equation}
\frac{D \rho}{Dt}  = -\rho \nabla \cdot \Ub = 
    \frac{1}{p_\rho}
    \left( \frac{D p}{Dt} - p_T \frac{D T}{Dt}
                          - \sum_k p_{X_k} {\omegadot}_k \right) \enskip ,
\end{equation}
with $p_\rho = \left.\partial p/\partial \rho\right|_{X_k,T}$,
$p_{X_k} = \left.\partial p/\partial X_k \right|_{T,\rho,(X_j,j\ne k)}$,
and $p_T = \left.\partial p/\partial T\right|_{\rho,X_k}$.

\subsection{Using the Enthalpy Equation}

Now writing $h = h(p,T,X_k)$ and expanding $Dh/Dt$ and using the enthalpy evolution equation:
\begin{equation}
\rho \frac{D h}{Dt}  = \rho \left( h_p \frac{D p}{Dt} + c_p \frac{D T}{Dt} + \sum_k h_{X_k} \frac{D X_k}{Dt} \right)
                     = \frac{D p}{D t} + \rho \Hnuc
\end{equation}
we can express $DT/Dt$ in terms of 
\begin{equation}
\frac{DT}{Dt} = \frac{1}{\rho c_p} \left( (1 - \rho h_p) \frac{D p}{D t}
- \sum_k \rho \xi_k \omegadot_k + \rho \Hnuc \right) \enskip , \label{eq:dTdt}
\end{equation}
where $c_p = \left.\partial h/\partial T\right|_{p,X_k}$ is the
specific heat at constant pressure,
$\xi_k = \left.\partial h/\partial X_k \right|_{p,T}$,
and $h_p = \left.\partial h/\partial p\right|_{T,X_k}.$

We could then write,
\begin{eqnarray*}
\nabla \cdot \ubold &=& \frac{1}{\rho p_\rho} \left(
- \frac{D p}{D t} + \frac{p_T}{\rho c_p}
  \left( (1 - \rho h_p) \frac{D p}{D t} - \rho \sum_k \xi_k \omegadot_k + \rho \Hnuc \right)
+ \sum_k p_{X_k} \omegadot_k \right)  \\
                 &=& \frac{1}{\rho p_\rho}
  \left( \frac{p_T}{\rho c_p}(1  - \rho h_p) - 1 \right) \frac{D p}{D t}
 + \frac{1}{\rho p_\rho} \left(
  \frac{p_T}{\rho c_p} (\rho \Hnuc - \rho \sum_k \xi_k   \omegadot_k)
                               + \sum_k p_{X_k} \omegadot_k \right) \enskip .
\end{eqnarray*}
When we derived this expression we explicitly retained the dependence of $h$ on $p$,
as shown by the presence of the $h_p$ term.

Then, replacing $p$ by $p_0(r)$, $Dp/Dt$ becomes $\ubold \cdot
\nabla p_0$, and the divergence constraint can be written
\begin{equation}
\nabla \cdot \ubold + \alpha \ubold \cdot \nabla p_0 =
\frac{1}{\rho p_\rho} \left(
   \frac{p_T}{\rho c_p} \left(
  - \sum_k\rho  \xi_k \omegadot_k + \rho \Hnuc \right)
 + \sum_k p_{X_k} \omegadot_k \right)  \equiv \tilde{S} \enskip \label{eq:full_divu_constraint} ,
\end{equation}
where we define
\begin{equation}
\alpha(\rho,T) \equiv - \left( \frac{(1 - \rho h_p )p_T - \rho c_p}{\rho^2
  c_p p_\rho} \right) \enskip . \label{eq:alphadef}
\end{equation}

\subsection{Using the Energy Equation}

Now writing $e = e(p,T,X_k)$ and expanding $De/Dt$ and using the energy evolution equation:
\begin{equation}
\rho \frac{D e}{Dt}  = \rho \left( e_p \frac{D p}{Dt} + c_v \frac{D T}{Dt} + \sum_k e_{X_k} \frac{D X_k}{Dt} \right)
                     = -p \nabla \cdot \Ub + \rho \Hnuc
\end{equation}
we can express $DT/Dt$ in terms of 
\begin{eqnarray*}
\frac{DT}{Dt} &=& \frac{1}{\rho c_v} \left(
                  -p \nabla \cdot \Ub + \rho \Hnuc
                 - \rho e_p \frac{D p}{Dt}
                 - \rho \sum_k e_{X_k} \omegadot_k   \right)
\end{eqnarray*}
where $c_v = \left.\partial e/\partial T\right|_{p,X_k}$ is the
specific heat at constant volume,
and $e_p = \left.\partial e/\partial p\right|_{T,X_k}.$
Then
\begin{equation}
-\rho \nabla \cdot \Ub = 
    \frac{1}{p_\rho} \left( \frac{D p}{Dt} 
    - \frac{p_T}{\rho c_v} \left(
                 -p \nabla \cdot \Ub + \rho \Hnuc
                 - \rho e_p \frac{D p}{Dt}
                 - \rho \sum_k e_{X_k} \omegadot_k   \right)
    - \sum_k p_{X_k} {\omegadot}_k \right) \enskip ,
\end{equation}
which leads to 
\begin{equation}
\left(-\rho -  \frac{p p_T}{\rho c_v p_\rho} \right) \nabla \cdot \Ub = 
    \frac{1}{p_\rho} \left( (1 + \frac{e_p p_T}{c_v}) \frac{D p}{Dt} 
                           - \frac{p_T}{\rho c_v} \left(
                                        \rho \Hnuc
                                        - \rho \sum_k e_{X_k} \omegadot_k   \right)
    - \sum_k p_{X_k} {\omegadot}_k \right) \enskip ,
\end{equation}

\section{Comparison of Constraints}

If we set $\omegadot_k = \Hnuc = 0$ for simplicity, then the constraint as derived using $h$ can be written 
\begin{equation}
\nabla \cdot \ubold
- \left( \frac{(1 - \rho h_p )p_T - \rho c_p}{\rho^2
  c_p p_\rho} \right) \frac{D p_0}{D t} = 0
\end{equation}
and the constraint derived using $e$ can be written
\begin{equation}
\left(\rho p_\rho + \frac{p p_T}{\rho c_v} \right) \nabla \cdot \Ub +
      (1 + \frac{e_p p_T}{c_v}) \frac{D p}{Dt} = 0 \enskip .
\end{equation}
